---
title: "p8105_hw6_iah2124"
author: "Iris Hart"
date: "2024-12-02"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(broom)
library(ggplot2)
library(rnoaa)
library(tidyr)
library(modelr)
```

## Problem 1

Loading the weather data
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

Set up the bootstrapping process
```{r}
set.seed(123)  

n_bootstrap <- 5000
```

Creating a function to perform regression and extract required metrics
```{r}
bootstrap_results <- replicate(n_bootstrap, {
  sample_data <- weather_df %>% sample_frac(replace = TRUE)
  
  lm_fit <- lm(tmax ~ tmin, data = sample_data)
  
  r_squared <- glance(lm_fit)$r.squared
  
  coef_values <- tidy(lm_fit)
  
  beta_0 <- coef_values$estimate[1]
  beta_1 <- coef_values$estimate[2]
  log_beta_prod <- log(beta_0 * beta_1)
  
  c(r_squared = r_squared, log_beta_prod = log_beta_prod)
}, simplify = TRUE)

bootstrap_df <- as.data.frame(t(bootstrap_results))
colnames(bootstrap_df) <- c("r_squared", "log_beta_prod")
```

Plot the distribution of r_squared
```{r}
ggplot(bootstrap_df, aes(x = r_squared)) +
  geom_density(fill = "skyblue", alpha = 0.5) +
  labs(title = "Distribution of Bootstrap Estimates for r^2", x = "r^2", y = "Density")
```
The ditribution of r_squared is approximately normally distributed with a 95% confidence interval of 0.8945701 ,  0.9271042. 

Plot the distribution of log(β0 * β1)
```{r}
ggplot(bootstrap_df, aes(x = log_beta_prod)) +
  geom_density(fill = "lightgreen", alpha = 0.5) +
  labs(title = "Distribution of Bootstrap Estimates for log(β0 * β1)", x = "log(β0 * β1)", y = "Density")
```
The distribution of log(β0 * β1) is approximately normally distributed with a 95% confidence interval of 1.964166 ,  2.058364. 


Calculate 95% Confidence Intervals
```{r}
ci_r_squared <- quantile(bootstrap_df$r_squared, c(0.025, 0.975))
ci_log_beta_prod <- quantile(bootstrap_df$log_beta_prod, c(0.025, 0.975))

cat("95% Confidence Interval for r^2: [", ci_r_squared[1], ", ", ci_r_squared[2], "]\n")
cat("95% Confidence Interval for log(β0 * β1): [", ci_log_beta_prod[1], ", ", ci_log_beta_prod[2], "]\n")
```

## Problem 2
Load data and filter out city states and races unneeded 
```{r}
homicide_df <- read_csv("homicide-data.csv") |>

  mutate(city_state = paste(city, state, sep = ",")) |>  

  filter(!(city_state %in% c("Dallas,TX", "Phoenix,AZ", "Kansas City,MO", "Tulsa,AL"))) |> 

  mutate(
    solved = case_when(
      disposition == "Closed without arrest" ~ 1,
      disposition == "Open/No arrest" ~ 0, 
      disposition == "Closed by arrest" ~ 1,
      TRUE ~ NA_real_  
    )
  ) |> 

  filter(victim_race %in% c("White", "Black")) |>  

  mutate(
    victim_age = as.numeric(victim_age)  
  )
```

```{r}
view(homicide_df)
```

Calculating the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed.
```{r}

baltimore_df <- homicide_df |>
  filter(city_state == "Baltimore,MD")

baltimore_reg <- glm(
  solved ~ victim_age + victim_sex + victim_race, 
  data = baltimore_df, 
  family = binomial()  
)

baltimore_OR <- broom::tidy(baltimore_reg) |>
  mutate(
    OR = exp(estimate), 
    CI_upper = exp(estimate + 1.96 * std.error),  
    CI_lower = exp(estimate - 1.96 * std.error)   
  ) |>
  filter(term == "victim_sexMale") |>
  select(OR, CI_lower, CI_upper)  

baltimore_OR
```

Calculating ORs for each city listed 
```{r}
allcity_ORresults <- homicide_df |>
  nest(data = -city_state) |>
  mutate(
    model = map(data, ~ glm(solved ~ victim_age + victim_sex + victim_race, data = .x, family = binomial())), 
    tidy_model = map(model, broom::tidy)
  ) |>
  select(-model, -data) |>
  unnest(cols = tidy_model) |>
  mutate(
    OR = exp(estimate), 
    CI_upper = exp(estimate + 1.96 * std.error),  
    CI_lower = exp(estimate - 1.96 * std.error)   
  ) |>
  filter(term == "victim_sexMale") |>
  select(city_state, OR, CI_upper, CI_lower)

allcity_ORresults |>
  knitr::kable(digits = 4)
```

Creating a plot to display OR data
```{r}
ggplot(allcity_ORresults, aes(x = reorder(city_state, OR), y = OR, ymin = CI_lower, ymax = CI_upper)) +
  geom_pointrange() +
  labs(
    title = "Estimated Odds Ratios for Solving Homicides: Male vs Female Victims",
    x = "City, State",
    y = "Odds Ratio (OR)",
    caption = "Whiskers represent 95% confidence intervals"
  ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 8),  
    axis.title = element_text(size = 10),  
    axis.text.x = element_text(size = 7, angle = 45, hjust = 1)  
  )
```
The estimated odds ratios for solving homicides for male vs. female victims is typically less than 1 for most cities. This indicates that more male victims are less likely compared to female victims to have a resolved homicide case. However, in Stockton, CA, Minneapolis, MN, and Fresno, CA, male victims are more likely to have a resolved homicide case than females. 

## Problem 3
Loading and cleaning birthweight data
```{r}
birthweight_df = read_csv("birthweight.csv") |> 
  mutate(babysex = as.factor(babysex),
         frace = as.factor(frace),
         malform = as.factor(malform),
         mrace = as.factor(mrace)
         )

view(birthweight_df)
```

View the data in a visual layout
```{r}
ggplot(birthweight_df, aes(x = bwt, y = blength)) + 
  geom_point(color = "coral") +   
  theme_minimal()
```


Propose the hypothetical fit model 
```{r}
hypotheticalrg_model <- lm(bwt ~ blength + bhead + babysex + gaweeks + delwt + fincome + 
                            smoken + wtgain + momage + mheight + parity, data = birthweight_df)

birthweight_df |> 
  add_predictions(hypotheticalrg_model) |>
  add_residuals(hypotheticalrg_model) |>
  ggplot(aes(x = pred, y = resid)) + 
  geom_point() + 
  geom_smooth(method = "lm", color = "lightgreen") + 
  theme_minimal() +
  labs(x = "Fitted Values", y = "Residuals")

```
### Model explanation ###

Main effects fitted model
```{r}
maineffects_model <- lm(bwt ~ blength + gaweeks, data = birthweight_df)

birthweight_df |>
  add_predictions(maineffects_model) |>
  add_residuals(maineffects_model) |>
  ggplot(aes(x = pred, y = resid)) + 
  geom_point() + 
  geom_smooth(method = "lm", color = "tomato") + 
  theme_minimal() +
  labs(x = "Fitted Values", y = "Residuals")
```
Interaction fitted model
```{r}
interaction_model <- lm(bwt ~ bhead * blength * babysex + gaweeks + delwt + smoken + 
                          wtgain + momage + mheight, data = birthweight_df)

birthweight_df |>
  add_predictions(interaction_model) |>
  add_residuals(interaction_model) |>
  ggplot(aes(x = pred, y = resid)) + 
  geom_point() + 
  geom_smooth(method = "lm", color = "orchid") + 
  theme_minimal() +
  labs(x = "Fitted Values", y = "Residuals")
```

Making comparisons between models
```{r}
cv_df <- crossv_mc(birthweight_df, 100) |> 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )

cv_res_df <- cv_df |> 
  mutate(
    hypotheticalrg_model = map(train, \(x) lm(bwt ~ blength + bhead + gaweeks + delwt + fincome + 
                                              smoken + wtgain + momage + mheight + parity, data = x)),
    maineffects_model = map(train, \(x) lm(bwt ~ blength + gaweeks, data = x)),
    interaction_model = map(train, \(x) lm(bwt ~ bhead * blength * babysex + gaweeks + delwt + smoken + 
                                             wtgain + momage + mheight, data = x))
  ) |> 
  mutate(
    rmse_hypothetical = map2_dbl(hypotheticalrg_model, test, rmse),
    rmse_main_effects = map2_dbl(maineffects_model, test, rmse),
    rmse_interaction = map2_dbl(interaction_model, test, rmse)
  )
```

Plotting comparisons 
